---
title: "Dynamic data visualization with R project"
author: "Otis Cooper and Stein Dijkstra"
date: "2023-01-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Goal

We have issues with the current landscape of rail planning apps
this is a first stage prototype within the scope of France
Choice of what to visualize
-choice to show stations
-departures
-prices
-Items
We believe that the project is feasible and can have benefits over current apps but more features need to be added.


## Key features

-Search your nearest station
-Updated Departures at each reload
-Full screen map
-Interactivity in plots and titles

## Potential new features

Use map to access departures and prices related to that station
Scheduling -> try to find route between locations with travelling time

## Known bugs
For some addresses, such as "0" pages is known to crash

## Online application

The application can be found at : 
https://stein-dijkstra.shinyapps.io/RailroadAid_Otis_Stein/

The code can be found at : 
https://github.com/Trebry/ENSAI_Data_visualization

## Documentation

```{r Load packages, message=FALSE, warning=FALSE, include=FALSE}
library(shiny)
library(shinyWidgets)
library(shinydashboard)
library(shinythemes)
library(colourpicker)
library(leaflet)
library(sf)
library(ggmap)
library(tidyverse)
library(geojsonR)
library(tidytransit)
```

```{r Load data, message=FALSE, warning=FALSE, include=FALSE}
# Download data using API
routeinfo <- read_delim("https://ressources.data.sncf.com/api/v2/catalog/datasets/tarifs-tgv-inoui-ouigo/exports/csv?limit=-1&offset=0&timezone=UTC",delim=";")
lost_found <- read_delim("https://ressources.data.sncf.com/api/v2/catalog/datasets/objets-trouves-restitution/exports/csv?order_by%3Ddate%20asc&limit=100&offset=0&timezone=UTC",delim = ";", escape_double = FALSE, trim_ws = TRUE)
gtfs <- read_gtfs("https://eu.ftp.opendatasoft.com/sncf/gtfs/export_gtfs_voyages.zip")

# Loading data that shouldn't change
stations_raw <- read_delim("mess/referentiel-gares-voyageurs.csv",delim = ";", escape_double = FALSE, trim_ws = TRUE)
horaires_gare<- read_delim("mess/horaires-des-gares1.csv",delim = ";", escape_double = FALSE, trim_ws = TRUE)
```

### Find a station
This tab shows all the different stations in France> Stations offering all the different services offered by SNCF are indicated by green arrows, smaller stations are indicated by orange arrows, and the smallest stations are indicated by red arrows.

When you click on a station you can see the name of the station and also it's opening hours is available.

In the shinny app you can also search for an address, and the app will pinpoint you to the closet station.

```{r Load map, echo=FALSE, message=FALSE, warning=FALSE}
# Tables used for the tab : Stations in France
horaires_gare <- horaires_gare %>%
  pivot_wider(names_from = 'Jour de la semaine', values_from = 'Horaire en jour normal', values_fn = list)
horaires_gare <- replace_na(horaires_gare, replace=list(rep("Not available", ncol(horaires_gare))))

stations_raw <- stations_raw %>% 
  left_join(horaires_gare,
            by = c('Code UIC'='UIC')) 

# Create map showing the various stations in France
  leaflet(data=stations_raw) %>%
    setView(lng = 2.2137, lat = 46.2276, zoom = 6) |>  addTiles() |> 
    addAwesomeMarkers(data = stations_raw[stations_raw$`Niveau de service` == '3',], 
                     ~Longitude, ~Latitude, icon = awesomeIcons(
                       icon = 'info-sign',
                       markerColor = 'green'
                     ), 
                     clusterOptions = markerClusterOptions(),
                     popup=~paste("<b>",`Intitulé plateforme`,
                                  `Code postal`,
                                  "<br>Horiaires d'ouverture : </b>",
                                  "<br>Monday ;", `Lundi`,
                                  "<br>Tuesday :", `Mardi`, 
                                  "<br>Wednesday :", `Mercredi`, 
                                  "<br>Thursday :", `Jeudi`, 
                                  "<br>Friday :", `Vendredi`, 
                                  "<br>Saturday :", `Samedi`,
                                  "<br>Sunday :", `Dimanche`)) |> 
    addAwesomeMarkers(data = stations_raw[stations_raw$`Niveau de service` == '2',], 
                   ~Longitude, ~Latitude, icon = awesomeIcons(
                     icon = 'info-sign',
                     markerColor = 'orange'
                   ), 
                   clusterOptions = markerClusterOptions(),
                   popup=~paste("<b>",`Intitulé plateforme`,
                                `Code postal`,
                                "<br>Horiaires d'ouverture : </b>",
                                "<br>Monday ;", `Lundi`,
                                "<br>Tuesday :", `Mardi`, 
                                "<br>Wednesday :", `Mercredi`, 
                                "<br>Thursday :", `Jeudi`, 
                                "<br>Friday :", `Vendredi`, 
                                "<br>Saturday :", `Samedi`,
                                "<br>Sunday :", `Dimanche`)) |> 
    addAwesomeMarkers(data = stations_raw[stations_raw$`Niveau de service` == '1',], 
                   ~Longitude, ~Latitude, icon = awesomeIcons(
                     icon = 'info-sign',
                     markerColor = 'red'
                   ),  
                   clusterOptions = markerClusterOptions(),
                   popup=~paste("<b>",`Intitulé plateforme`,
                                `Code postal`,"</b>"))
```

### Departures

This tab shows all the departures from a selected station. You can also search the various routes leaving the station using the search bar on the right.

Example of data shown for a depture from Rennes :

```{r Show departures, echo=FALSE, message=FALSE, warning=FALSE}
stop_ids <- gtfs$stops %>% select(stop_id,stop_name) 
departures <- stop_ids %>% 
  inner_join(gtfs$stop_times %>% 
               select(trip_id, arrival_time, 
                      departure_time, stop_id), 
             by = "stop_id")

departures <- departures %>% 
  left_join(gtfs$trips %>% 
              select(trip_id, route_id, 
                     service_id, trip_headsign), 
            by = "trip_id") 
departures <- departures %>% 
  left_join(gtfs$routes %>% 
              select(route_id, 
                     route_long_name,route_desc), 
            by = "route_id")
departures <- departures %>% 
  left_join(gtfs$calendar_date %>% 
              select(service_id, 
                     date), 
            by = "service_id")

  departures %>% filter(stop_name == "Rennes") %>% 
    select(stop_name,date,
           arrival_time,
           departure_time,
           trip_headsign,
           route_long_name) %>%
    arrange(date,arrival_time)

```

### Prices

This tab allows you to select a departure station, the class, the minimum and maximum price. When this three criteria have been chosen, the app will display a histogram of the train ticket prices and also a table of the various ticket prices.

An example of data shown for a depture from rennes, in class 1 at minimum price : 

```{r Show prices, echo=FALSE, message=FALSE, warning=FALSE}

fromLocation = "RENNES"
class = "1"
price = "minimum"
bins = 30

  x    <-   routeinfo %>% filter(`origine`== (if (fromLocation=="") "RENNES" else fromLocation)) %>% 
    filter(`classe`==class) %>% pull( (if (price=="minimum") `prix_min` else `prix_max`))
  
  bins <- seq(min(x), max(x), length.out = bins + 1)
  
  # draw the histogram with the specified number of bins
  hist(x, breaks = bins, border = 'white',
       xlab = '',
       main = paste("Histogram of",price,"ticket prices from",fromLocation,"in class",class))
  
  # Display table
    routeinfo %>% filter(`origine`==fromLocation) %>% 
    filter(`classe`==class)
  
```

### Lost items

This tab shows the various lost items in stations and in trains. The panels is composed of 3 tabs, a tab showing a bar chart of the number of items lost per category, a pie chart showing the percentage of each category of lost items, and final a table showing the details of the lost items>

On the left of the tab, we can choose which lost items to show.

Note this tab, for performance reason only shows 100 lost items.

Example of data shown for lost items :

```{r Show lost items, echo=FALSE, message=FALSE, warning=FALSE}
  ggplot(data = lost_found
    , aes(x=factor(1), fill=gc_obo_type_c))+
    geom_bar(width = 1)+
    coord_polar("y")+
    theme(axis.title.x=element_blank(),
          axis.text.x=element_blank(),
          axis.ticks.x=element_blank(),
          axis.title.y=element_blank(),
          axis.text.y=element_blank(),
          axis.ticks.y=element_blank()) +
    guides(fill=guide_legend(title="Lost items"))

  ggplot(data = lost_found) +
    geom_bar(mapping = aes(x = gc_obo_type_c, fill = gc_obo_type_c)) + 
    coord_flip() +
    theme(axis.title.y=element_blank(),
          legend.position = "none")

```
